---
title: "STAT 3301"
author: "A MINI - Project 3"
date: "Yoojin Jeong (jeong.358)"
output: html_document

  
---

<!-- These are some LaTeX macros for certain symbols -->
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bfb}{\mbox{\boldmath $\beta$}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\bY}{\mathbf{Y}}

```{r message=FALSE,echo=FALSE}
library(readr)
library(tidyverse)
library(alr4)
library(broom)
```
\


\large
#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions carefully.
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. 
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
- formatting all plots to be appropriately sized, with appropriate axis labels.
- only including R code that is necessary to answer the questions below.
- only including R output that is necessary to answer the questions below (avoiding lengthy output).
- providing short written answers explaining your work, and writing in complete sentences.

**Introduction**

This is a continuation of the project 2.

Do time young children remain at the lunch table and type of food they take help predict how much they eat in terms calories? **$n=126$** toddlers were observed over several months at a nursery school with a standard food type and two newly introduced food types in a school. The data set "CaloriesProject2" for this project is available on Carmen. 

The variables are:

| Variable Name | Description |
| ------------- | --------------------------------------------------------------------- |
| `time` | The number of **minutes** a child spent at the table when lunch was served. |
| `type` | **What type of food** children eat at the lunch time. There are three different type of food: one is the standard food type, and other two are the newly introduced food types (`std`, `New1`, `New2`). |
| `calories` | **number of calories** the child consumed during lunch, calculated from careful observation of what the child ate on a Friday |


In our project 2, we have considered the three different models to predict calories with different predictors and different combinations. The three potential models:

M1. Assumes the relationship between `coloreis` and `time` is 
exactly the same no matter the types of food (`calories ~ time`).
This is the one we considered in our mid exam 1 with some modification 

M2. Assumes parallel regression lines (`calories~time+type`). 
The effect of time on calories is the same for all food types and vice versa. 

M3: Assumes different regression relationships (`calories~time*type`). 
The effect of time on calories is not the same for all food types and vice versa.

We are interested in applying the k-fold cross-validation to select a best models out the three above considered models.

\vspace{1cm}
**Part (i)**
\vspace{1cm}

Consider $k=5$ with the following data splitting:
split the data into five roughly equal groups. There are $n = 126$ samples in the data set, and so we can split the data into first four groups of 25 data points each and the last group of 26 data points. 

**Please use the seed number `33013301` for you sampling**

(1) Compute cross validated average of the prediction errors for each model:
\[
CV(M_j) = \frac{1}{126}\sum_{i=1}^{126} (y_i - \bx_{i, M_j}^T \hat{\bfb}_{M_j}^{-O(i)})^2
\]
for $j=1,2,3$ and with 

$y_i$= i^th^ response value of the original data.

$x_i^T$= is the predictor values of the i^th^ observation of the original data.

$\hat{\bfb}_{M_j}^{-k} = ({\bX_{M_j}^{-k}}^T \bX_{M_j}^{-k})^{-1} {\bX_{M_j}^{-k}}^T \bY^{-k}$
be the estimated regression coefficients for model $M_j$ based on the training data (i.e., the data
without set $k$).

where an individual case $i$, $O(i)$, let $O(i) \in \{1, 2, \ldots, 5\}$ specify to which set the case
belongs.

```{r}
cal2 = read_csv("CaloriesProject2.csv")
set.seed(33013301)
calories = cal2 %>% mutate(fold = sample(head(rep(1:5, 26), 126)))

M1 = M2 = M3 = vector("list", 5)
for(k in 1:5) {
  M1[[k]] = lm(calories~time, data = calories %>% filter(fold != k))
  M2[[k]] = lm(calories~time+type, data = calories %>% filter(fold != k))
  M3[[k]] = lm(calories~time*type, data = calories %>% filter(fold != k))
}

preds.M1 = preds.M2 = preds.M3 = rep(NA, 126)

for(i in 1:126) {
  foldi = calories$fold[i]
  preds.M1[i] = predict(M1[[foldi]], newdata = calories[i,])
  preds.M2[i] = predict(M2[[foldi]], newdata = calories[i,])
  preds.M3[i] = predict(M3[[foldi]], newdata = calories[i,])
}

mean((calories$calories - preds.M1)^2)
mean((calories$calories - preds.M2)^2)
mean((calories$calories - preds.M3)^2)
```


(2) Select the best model based on the cross validated average of the 
prediction errors computed in question (1).
\vspace{1cm}

Among these three models, M3 seems to be the best model because the cross validated average of the
prediction error is the smallest.

